\section{Hardware Implementation Results}

This chapter summarizes the LUT, FF, DSP, BRAM, clock frequency, power, and normalized metrics obtained from Vivado synthesis and implementation.

The current mainline device has already switched to \texttt{XCZU4EV-SFVC784-2I}, with the default target period set to \texttt{3.333 ns}. So far, all five custom RTL architectures have completed ZU4EV post-route data collection, and among them \texttt{fir\_pipe\_systolic} is the only version that achieves positive timing margin at this target period, with an estimated \texttt{Fmax = 459.348 MHz} while also keeping the lowest \texttt{energy/sample}.

\texttt{fir\_l2\_polyphase} and \texttt{fir\_l3\_polyphase} prove the implementability of parallel polyphase datapaths on a larger device: the former reaches about \texttt{278.668 MS/s}, and the latter about \texttt{381.388 MS/s}. However, from the current results, \texttt{L=3} still does not surpass the overall synthesis quality of the high-quality scalar pipeline, so the hero architecture on the ZU4EV platform remains \texttt{fir\_pipe\_systolic}. The complete data uses \texttt{data/impl\_results.csv} in the repository as the single source of truth.

This round also filled in the \texttt{board-shell scope} results, which fold \texttt{PS + AXI DMA + FIR shell + bare-metal harness} into the implementation statistics as a complete system. Under this framing, both \texttt{zu4ev\_fir\_pipe\_systolic\_top} and \texttt{zu4ev\_fir\_vendor\_top} completed bitstream generation, `.xsa` export, and on-board validation. The former is estimated at \texttt{Fmax = 347.826 MHz}, \texttt{LUT = 20253}, and \texttt{energy/sample = 8.542 nJ}; the latter at \texttt{Fmax = 347.102 MHz}, \texttt{LUT = 8856}, and \texttt{energy/sample = 8.243 nJ}. This shows that in the final system-level comparison, the vendor FIR IP is slightly better in resource use and energy efficiency, while the custom system shell keeps almost the same operating frequency.

To make these results more interpretable, this round also added normalized efficiency metrics. Under kernel scope, for example, \texttt{fir\_pipe\_systolic} reaches about \texttt{3.480} in \texttt{throughput/DSP}, clearly exceeding \texttt{fir\_symm\_base} and \texttt{fir\_l2\_polyphase}; meanwhile, \texttt{fir\_l2\_polyphase} shows very high \texttt{throughput/kLUT}, indicating strong LUT utilization efficiency but a relatively high DSP cost. Under the system-shell framing, the ordering changes: \texttt{vendor} and \texttt{custom} run at nearly identical frequency, but vendor achieves higher \texttt{throughput/kLUT} and lower \texttt{energy/sample}.

The current power framing comes from the vectorless routed \texttt{report\_power} estimate, with confidence level \texttt{Medium}, so these numbers should be interpreted as relative-comparison data rather than final sign-off power. For this project, the more valuable insight is the layered power breakdown: under board-shell scope, \texttt{PS8} is nearly fixed at \texttt{2.228 W}, while the real FIR-structure difference appears in \texttt{fir\_shell\_0}, which is about \texttt{0.238 W} for the custom system shell and about \texttt{0.132 W} for the vendor system shell.

The timing analysis also gives an important conclusion. The current board-shell worst path is no longer inside the FIR core DSP MAC itself, but on the path from FIR output through \texttt{round/saturate} into the output FIFO BRAM. In the custom system shell, this path has \texttt{Data Path Delay = 2.859 ns}, including about \texttt{1.348 ns} logic delay and about \texttt{1.511 ns} routing delay, so routing accounts for more than half of the path. This indicates that once the FIR kernel itself is highly pipelined, the system-level performance bottleneck naturally shifts to the output boundary and local interconnect rather than staying on the multiplier.
